Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2014-03-28 parts/sa.tex
global proofread

We consider in this section a PH model $\PH = (\PHs; \PHl; \PHa^{\langle 2 \rangle})$ with $2$ classes of priorities
that respects all conditions (\ref{cr:bounded} \& \ref{cr:compcs}) of \pref{ssec:hypothesis}.

@@

\begin{definition}[$\concr: \OS \to \powerset(\Sce)$]\label{def:concr}
Given $\w\in\OS$, $\concr(\w)$ is the set of minimal scenarios concretizing $\w$ in the
context $\ctx$. It is defined as the largest set satisfying the following conditions:
\begin{enumerate}[(i)]
\item $\forall\delta\in\concr(\w), \exists s\subseteq\ctx: \delta\in\Sce(s)$
\item $\forall\delta\in\concr(\w),
  \exists \phi:\indexes{\w}\to\indexes{\delta},
    (\forall n,m\in\indexes{\w}, n<m \Leftrightarrow \phi(n)\leq\phi(m)):$
\\\hspace*{2cm}$\forall n\in\indexes{\w}, \PHbounce(\w_n) \in \ctx\play\delta_{1..\phi(n)}$
\item $\forall\delta,\delta'\in\concr(\w)
				\card{\delta}\leq\card{\delta'} \Rightarrow
					\delta\neq\delta'_{1..\card{\delta}}$.
\end{enumerate}
\end{definition}
\begin{definition}[$\uconcr: \OS \rightarrow \powerset(\Sce)$]
\label{def:uconcr}
  \[
  \uconcr(\w) \DEF
  \begin{cases}
    \concr(\w) & \text{if } \forall s \in \PHl, s \subseteq \ctx, \exists \delta \in \concr(\w), \delta \in \Sce(s) \\
    \emptyset & \text{else.}
  \end{cases}
  \]
\end{definition}
% 
\begin{lemma}
\label{lem:uconcr-ctx}
  $\ctx \subseteq \ctx' \wedge \muconcr_{\ctx'}(\w) \neq \emptyset \Longrightarrow \muconcr_{\ctx}(\w) \neq \emptyset$.
\end{lemma}

@@

\begin{example}
  Let $\PH' = (\PHs, \PHl, \PHh'^{\langle 1 \rangle})$ be the “flattened” version of the PH in \pref{fig:ph-livelock}, that is: $\PHh'^{\langle 1 \rangle} = \PHh^{(1)} \cup \PHh^{(2)}$,
  which is equivalent to a PH in the semantics without priorities.
  Due to spurious behaviours inherent to the cooperative sorts in this semantics, 
  the original under-approximation developed in~\cite{PMR12-MSCS} concludes that $c_1$ is reachable in $\PH'$ from $\ctx = \PHstate{a_1, b_0, c_0, ab_{10}}$.
  %this is due to an over-approximation of the behaviours inherent to cooperative sorts in the semantics without priorities.

@@

  However, if $\PHhit{a_0}{b_0}{b_1}$ and $\PHhit{b_0}{a_0}{a_1}$ are replaced by
  $\PHhit{a_0}{a_0}{a_1}$ and $\PHhit{b_0}{b_0}{b_1}$,
  then \pref{th:approxinf} concludes that $c_1$ is reachable from $\ctx$.

@@

    (b0) edge (b10) edge (b00)
    (b10) edge (b10s)
    (b00) edge (b00s)
    (b1) edge (b01) edge (b11)
    (b01) edge (b01s)
    (b01s) edge (a0)
    (b11) edge (b11s)
    ;
    \end{tikzpicture}
  \caption{
  \label{fig:sa-livelock}
    The graph of local causality of the PH model in \pref{fig:ph-livelock}.
    Rectangular nodes containing a single process are elements in $\Proc$,
    nodes containing a couple of processes are elements in $\Obj$
    and circle nodes are elements in $\Sol$.
    %The double arrow and the two surrounding nodes represent the only edge in $\Proc \times \Sol$.
    \pref{th:approxinf} is inconclusive on this example as edge $(ab_{11}, \{ a_1, b_1 \}) \in \Proc \times \Sol$ (here represented with a double line) is not coherent (\pref{def:coherent}).
    Indeed, $a_0 \in \Proc$ is a child of $\{ a_1, b_1 \}$, but $a_0 \neq a_1$ (and the same also stands for $b_0$).
  }
\end{figure}
\end{example}

@@

The reachability property studied so far concern a single process at a time.
However, we remark that the reachability of a partial or global state can be
addressed with the very same analysis by introducing a dedicated cooperative
sort in the scope of the PH semantics with $2$ classes of priorities.

@@

It is also possible to compute the reachability of a set of states $S \subseteq \PHl$ by creating several actions $\PHhit{\total_s}{\reach_0}{\reach_1}$ in $\PHh^{(2)}$ for each state $s \in S$.

@@

Let us define a sequence of objective $\w=\obj{a_i}{a_j}\concat\w'$ with
$a_i\neq a_j$ and a state $s\in L$ with $\get{s}{a}=a_i$.
One can remark that any scenario reaching $a_j$ necessarily includes one of the
bounce sequence in $\BS(\obj{a_i}{a_j})$, and in particular,
any minimal scenario reaching $a_j$ ends in a state where $a_j$ is present but
also the hitter of the last of one bounce sequence in $\BS(\obj{a_i}{a_j})$.
If such hitter $b_k$ is in a cooperative sort $b\in\cs$, it also means that one of the sub-state
in $\csState(b_k)$.
\pref{def:lastprocs} defines $\lastprocs(\obj{a_i}{a_j})$ as the set of set of
processes that may be present just after reaching $a_j$.

@@

\begin{definition}[$\lastprocs$]
\label{def:lastprocs}
Given an objective $\obj{a_i}{a_j}$, $\lastprocs:
\Obj\to\powerset(\powerset(\Proc))$ is
defined such that, $\forall lps\in\lastprocs(\obj{a_i}{a_j})$, 
$lps\in\powerset(\Proc)$,
\begin{enumerate}
\item $\{\hitter{\zeta_{\card\zeta}}\} \subset lps$;
\item $\forall b_j\in lps, b\in\cs, \exists ps\in\csState(b_j): ps\subset lps$;
\item $\forall b_j\in lps, b\in\cs, 
		\forall ps\in\csState(b_j),
			\exists lps'\in\lastprocs(\obj{a_i}{a_j}),
				ps\subset lps'$;
\item $\nexists lps'\in\lastprocs(\obj{a_i}{a_j}): lps'\subset lps \wedge
									lps'\neq lps$.
\end{enumerate}
\end{definition}

\begin{theorem}[Sequential under-approximation]
\label{thm:ordered-ua}
Given a context $\ctx$ and an objective sequence $\w =
P\concat\w'\in\OS$,
$\uconcr(P)\neq\emptyset \wedge
	\forall ps \in\lastprocs(P),
	\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset
	\Longrightarrow \uconcr(\w)\neq\emptyset$,
where $\ctx' = \ctx\Cap\procs(\mycwB{\ctx}{P})$.
\end{theorem}
\begin{proof}
If $\uconcr(P)\neq\emptyset$,
for all $s\in L, s\subset\ctx$,
there exists a scenario $\delta\in\uconcr(P)\cap\Sce(s)$;
from \pref{def:lastprocs} and proof of \pref{th:approxinf},
$\exists ps\in\lastprocs(P)$ such that
$(s\play\delta)\subset\ctx'\Cap ps$.
Hence, if $\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset$,
there exists a scenario $\delta'\in\muconcr_{\ctx'\Cap ps}(\w')$ such that
$\delta'\in\Sce(s\play\delta)$.
Hence, $\delta\concat\delta'$ is a scenario playable in $s$.
Therefore, for all $s\in L, s\subset\ctx$, there exists a scenario
concretizing $\w$.
Hence, $\uconcr(\w)\neq\emptyset$.
\end{proof}

>>>>

We consider in this section a PH model $\PH = (\PHs; \PHl; \PHa^{\langle 2 \rangle})$ with $2$ classes of priorities
that respects all Conditions (\ref{cr:bounded} \& \ref{cr:compcs}) of \pref{ssec:hypothesis}.

@@

\begin{definition}[$\concr: \OS \to \powerset(\Sce)$]\label{def:concr}
Given $\w\in\OS$, $\concr(\w)$ is the set of minimal scenarios concretizing $\w$ in the
context $\ctx$. It is defined as the largest set satisfying the following conditions:
\begin{enumerate}[(i)]
\item $\forall\delta\in\concr(\w), \exists s\subseteq\ctx, \delta\in\Sce(s)$
\item $\forall\delta\in\concr(\w),
  \exists \phi:\indexes{\w}\to\indexes{\delta},
    (\forall n,m\in\indexes{\w}, n<m \Leftrightarrow \phi(n)\leq\phi(m)),$
\\\hspace*{2cm}$\forall n\in\indexes{\w}, \PHbounce(\w_n) \in \ctx\play\delta_{1..\phi(n)}$
\item $\forall\delta,\delta'\in\concr(\w)
				\card{\delta}\leq\card{\delta'} \Rightarrow
					\delta\neq\delta'_{1..\card{\delta}}$.
\end{enumerate}
\end{definition}
\begin{definition}[$\uconcr: \OS \rightarrow \powerset(\Sce)$]
\label{def:uconcr}
  \[
  \uconcr(\w) \DEF
  \begin{cases}
    \concr(\w) & \text{if } \forall s \in \PHl, s \subseteq \ctx, \exists \delta \in \concr(\w), \delta \in \Sce(s) \\
    \emptyset & \text{else.}
  \end{cases}
  \]
\end{definition}
% 
\begin{lemma}
\label{lem:uconcr-ctx}
  $\ctx \subseteq \ctx' \wedge \muconcr_{\ctx'}(\w) \neq \emptyset \Longrightarrow \muconcr_{\ctx}(\w) \neq \emptyset$.
\end{lemma}

@@

\begin{example}
  Let $\PH' = (\PHs, \PHl, \PHh'^{\langle 1 \rangle})$ be the “merged” version of the PH in \pref{fig:ph-livelock}, that is: $\PHh'^{\langle 1 \rangle} = \PHh^{(1)} \cup \PHh^{(2)}$,
  which is how it would have been represented in the semantics without priorities.
  Due to spurious behaviours inherent to the cooperative sorts in this semantics, 
  the original under-approximation developed in~\cite{PMR12-MSCS} concludes that $c_1$ is reachable in $\PH'$ from $\ctx = \PHstate{a_1, b_0, c_0, ab_{10}}$.

@@

  However, if $\PHhit{a_0}{b_0}{b_1}$ and $\PHhit{b_0}{a_0}{a_1}$ are replaced by the auto-actions
  $\PHhit{a_0}{a_0}{a_1}$ and $\PHhit{b_0}{b_0}{b_1}$,
  then \pref{th:approxinf} concludes that $c_1$ is reachable from $\ctx$.

@@

    (b0) edge (b10) edge (b00)
    (b10) edge (b10s)
    (b00) edge (b00s)
    (b1) edge (b01) edge (b11)
    (b01) edge (b01s)
    (b01s) edge (a0)
    (b11) edge (b11s)
    ;
    \end{tikzpicture}
  \caption{
  \label{fig:sa-livelock}
    The graph of local causality of the PH model in \pref{fig:ph-livelock}
    for the objective $\w = \PHobj{c_0}{c_1}$ and the initial context $\ctx = \PHstate{a_1, b_0, c_0, ab_{10}}$.
    Rectangular nodes containing a single process are elements in $\Proc$,
    borderless nodes containing a couple of processes are elements in $\Obj$
    and circle nodes are elements in $\Sol$.
    %The double arrow and the two surrounding nodes represent the only edge in $\Proc \times \Sol$.
    \pref{th:approxinf} is inconclusive on this example as edge $(ab_{11}, \{ a_1, b_1 \}) \in \Proc \times \Sol$ (here represented with a double line) is not coherent (\pref{def:coherent}).
    Indeed, $a_0 \in \Proc$ is a child of $\{ a_1, b_1 \}$, but $a_0 \neq a_1$ (and the same also stands for $b_0$).
  }
\end{figure}
\end{example}

@@

The reachability property studied so far concerns a single process at a time.
However, we remark that the reachability of a partial or global state can be
addressed with the very same analysis by introducing a dedicated cooperative
sort in the scope of the PH semantics with $2$ classes of priorities.

@@

It is also possible to compute the reachability of a set of states $S \subseteq \PHl$ by creating several actions $\PHhit{\total_s}{\reach_0}{\reach_1}$ in $\PHh^{(2)}$ for each state $s \in S$,
or on a sub-state $s \in \PHsubl[\PHl]_S$ of a set of components $S \subset \PHs$ by adapting the sort $\total$ with: $\PHl_\total = \PHsubl[\PHl]_S$

@@

Let us define a sequence of objectives $\w=\obj{a_i}{a_j}\concat\w'$ with
$a_i\neq a_j$ and a state $s\in \PHl$ with $\get{s}{a}=a_i$.
One can remark that any scenario reaching $a_j$ necessarily includes one of the
bounce sequences in $\BS(\obj{a_i}{a_j})$, and, in particular,
any minimal scenario reaching $a_j$ ends in a state where $a_j$ is present but
also the hitter of the last bounce of one bounce sequence in $\BS(\obj{a_i}{a_j})$.
If such hitter $b_k$ is in a cooperative sort $b\in\cs$, it also means that one of the sub-states
in $\csState(b_k)$ is included in the ending state.
\pref{def:lastprocs} defines $\lastprocs(\obj{a_i}{a_j})$ as the set of set of
processes that may be present just after reaching $a_j$.

@@

\begin{definition}[$\lastprocs$]
\label{def:lastprocs}
Given an objective $\obj{a_i}{a_j}$, $\lastprocs:
\Obj\to\powerset(\powerset(\Proc))$ is
defined such that, $\forall lps\in\lastprocs(\obj{a_i}{a_j})$, 
$lps\in\powerset(\Proc)$,
\begin{enumerate}
\item $\{\hitter{\zeta_{\card\zeta}}\} \subset lps$;
\item $\forall b_j\in lps, b\in\cs, \exists ps\in\csState(b_j), ps\subset lps$;
\item $\forall b_j\in lps, b\in\cs, 
		\forall ps\in\csState(b_j),
			\exists lps'\in\lastprocs(\obj{a_i}{a_j}),
				ps\subset lps'$;
\item $\nexists lps'\in\lastprocs(\obj{a_i}{a_j}), lps'\subset lps \wedge
									lps'\neq lps$.
\end{enumerate}
\end{definition}

\begin{theorem}[Sequential under-approximation]
\label{thm:ordered-ua}
Given a context $\ctx$ and an objective sequence $\w =
P\concat\w'\in\OS$,
$\uconcr(P)\neq\emptyset \wedge
	\forall ps \in\lastprocs(P),
	\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset
	\Longrightarrow \uconcr(\w)\neq\emptyset$,
where $\ctx' = \ctx\Cap\procs(\mycwB{\ctx}{P})$.
\end{theorem}
\begin{proof}
If $\uconcr(P)\neq\emptyset$,
for all $s\in \PHl, s\subset\ctx$,
there exists a scenario $\delta\in\uconcr(P)\cap\Sce(s)$;
from \pref{def:lastprocs} and proof of \pref{th:approxinf},
$\exists ps\in\lastprocs(P)$ such that
$(s\play\delta)\subset\ctx'\Cap ps$.
Hence, if $\muconcr_{\ctx'\Cap ps}(\w')\neq\emptyset$,
there exists a scenario $\delta'\in\muconcr_{\ctx'\Cap ps}(\w')$ such that
$\delta'\in\Sce(s\play\delta)$.
Hence, $\delta\concat\delta'$ is a scenario playable in $s$.
Therefore, for all $s\in \PHl, s\subset\ctx$, there exists a scenario
concretizing $\w$.
Hence, $\uconcr(\w)\neq\emptyset$.
\end{proof}
